---
title: 'linearprog: An R Package for Inference in Linear and Quadratic Programs'
author: ''
output:
  pdf_document:
    toc: yes
    df_print: paged
  html_document:
    keep_md: yes
    toc: yes
    df_print: paged
  github_document:
    keep_html: yes
    toc: yes
    df_print: paged
bibliography: refs.bib
vignette: |
  %\VignetteIndexEntry{dkqs} 
  %\VignetteEncoding{UTF-8} 
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

```{r setup, include = FALSE}
require(linearprog)
require(data.table)
require(pander)
require(gurobi)
require(knitr)
require(kableExtra)
require(Momocs)
```

## Introduction
This package conducts inference on econometrics problems that can be 
represented by quadratic programming. This package uses the cone-tightening 
procedure of @dkqs2018 and the procedure of choosing the optimal value of tau by 
@kamat2019.

## Scope of the Vignette
This vignette is intended as a guide to using the `linearprog` package. Readers
may refer to section 4.2 of @dkqs2018 for details about the cone-tightening 
procedure and the supplemental appendix of @kamat2019.

## Installation and Requirements
`linearprog` can be installed from our GitHub repository via

```{r eval = FALSE}
devtools::install_github("conroylau/linearprog")
```

To use `linearprog`, one of the following packages for solving linear and
quadratic programs are required. There are four options for the packages:

i. Gurobi and the R package `gurobi` --- Gurobi can be downloaded from
[Gurobi Optimization](https://www.gurobi.com/). A Gurobi software license is 
required for this software, which can be obtained at no cost for academic
researchers. The instructions for installing `gurobi` on R can be found 
[here](https://cran.r-project.org/web/packages/prioritizr/vignettes/gurobi_installation.html#r-package-installation).

i. `limSolve` --- a free and open-source package available on CRAN. This can be
installed directly via the `install.packages` command in R.

i. IBM ILOG CPLEX Optimization Studio (CPLEX) and one of the R packages below 
--- CPLEX can be downloaded from 
[IBM](https://www.ibm.com/analytics/cplex-optimizer).
The R packages are available on CRAN. Note that users who are using OS X would 
need to install the packages on the command line to link the package to the 
correct CPLEX library. Both R packages are open-source packages. Any one of the
following packages can be used:
   i.  `Rcplex`  --- the instructions to install the R package can be found
   [here](https://cran.r-project.org/web/packages/Rcplex/INSTALL).
   
   i. `cplexAPI` --- the instructions to install the R package can be found
   [here](https://cran.r-project.org/web/packages/cplexAPI/INSTALL).

If no package is specified, one of the above packages will be automatically 
chosen from those that are available.

## Usage Demonstration 1: Missing Data Problem
### Background and Data {#data}
This missing data problem is a classical problem due to @manski1989, which is
demonstrated by the dataset below. The dataset contains 1,000 rows with 2 
columns where `D` is binary and `Y` takes 11 distinct values. 

The sample data is included with the `linearprog` package. The sample data 
has two columns and the following structure.
```{r, drawData-df}
library(linearprog)
knitr::kable(head(sampledata, n = 10))
```
The two variables that are used are as follows:

* `Y` is a multivariate discrete outcome variable taht takes value from 0 to 1 
with step size 0.1.
* `D` is a binary treatment where *D*~*i*~ = 1 means that *Y*~*i*~ is observed.

### Syntax  {#syntax}
The main command for applying the cone-tightening procedure in `linearprog` is 
called `dkqs_cone`.
```{r, syntax, eval = FALSE}
library(linearprog)
dkqs_cone(df = sampledata, 
          A_obs = A_obs_twom, 
          A_tgt = A_target, 
          func_obs = func_two_moment, 
          beta_tgt = 0.375, 
          bs_seed = 1,
          bs_num = 100,
          p_sig = 2,
          tau_input = tau,
          solver = gurobi)
```
where

* `df` refers to the data being used in the inference.
* `A_obs` refers to the "observed matrix" in the inference.
* `A_tgt` refers to the "target matrix" in the inference.
* `func_obs` refers to the function that generates the vector of observed beta.
* `beta_tgt` refers to the value of beta to be tested.
* `bs_seed` refers to the starting value of the seed in bootstrap.
* `bs_num` refers to the total number of bootstraps to be conducted.
* `p_sig` refers to the number of decimal places in the p-value.
* `tau_input` refers to the value of tau chosen by the user.
* `solver` refers to the solver used to solve the linear and quadratic
programs.

The following two sections explain how the required parameters and functions
can be constructed in order to apply the `dkqs_cone` function.

### Specifying the Functions {#functions}

The `dkqs_cone` function provides flexibility for users to specify the functions to generate the observed value of beta from the dataset. The 
requirement for the function provided by the user is as follows:

* It only takes one argument, which is the dataset.
* It only returns a numeric vector that corresponds to the vector for the 
observed beta. 

#### Full Information Approach

The following is an example of defining the function for the full information
approach:

```{r, syntax1}
func_full_info <- function(df){
  beta = NULL
  y_list = sort(unique(df[,"Y"]))
  n = dim(df)[1]
  yn = length(y_list)
  for (i in 1:yn){
    beta_i = sum((df[,"Y"] == y_list[i]) * (df[,"D"] == 1))/n
    beta = c(beta,c(beta_i))
  }
  beta = as.matrix(beta)
  return(beta)
}
```
The `func_full_info` function returns a vector of length *J* + 1 where element 
*i* of the vector refers to the probability that the corresponding value of 
*y*~*i*~ is observed.

#### Two Moments Approach

The following is an example of defining the function for the two moments 
approach:

```{r, syntax2}
func_two_moment <- function(df){
  beta = matrix(c(0,0), nrow = 2)
  n = dim(df)[1]
  beta[1] = sum(df[,"Y"] * df[,"D"])/n
  beta[2] = sum(df[,"D"])/n
  return(beta)
}
```
The `func_two_moment` function returns a vector with two elements that 
corresponds to the two moments.

### Specifying the parameters
As shown in the syntax section, the matrices `A_obs` and `A_tgt` have to be 
defined in order to use the function. To construct the two matrices, the
following parameters are needed:
```{r, results='markdown'}
N = dim(sampledata)[1]
J1 = length(unique(sampledata[,"Y"]))
yp = seq(0,1,1/(J1-1))
```

With the above quantities, the two matrices can be defined as follows:
```{r, results='markdown'}
A_obs_twom = matrix(c(rep(0,J1), yp, rep(0,J1), rep(1, J1)), nrow = 2,
                byrow = TRUE)
A_target = matrix(c(yp, yp), nrow = 1)
```

Lastly, the value of tau can be defined freely by the user as long as the
quadratic program is feasible. Here, we choose the value of tau based on the 
formula from the supplemental appendix in page 15 of @kamat2019:
```{r, results='markdown'}
tau = sqrt(log(N)/N)
```

### Output {#output}
The following are the output when the two moments approach is used with the
`gurobi` solver to test the hypothesis that `beta_tgt` is 0.375.

```{r, syntax.eval, eval = TRUE}
library(linearprog)
dkqs_cone(df = sampledata, 
         A_obs = A_obs_twom, 
         A_tgt = A_target, 
         func_obs = func_two_moment, 
         beta_tgt = 0.375, 
         bs_seed = 1,
         bs_num = 100,
         p_sig = 3,
         tau_input = tau,
         solver = "gurobi")
```

## Help, Feature Requests and Bug Reports
Please post an issue on the 
[GitHub repository](https://github.com/conroylau/linearprog/issues).

## References



 